Implementation and assessment of tools for integrating forest roads into projections of the cumulative effects of disturbance on wildlife

### 2024-05-15 analysis/paper/supplement.Rmd
fixing gdistance profiling and updating text to include it

The simple cost method for setting edge weights generally performed worse than the grade penalty and Hardy methods (compare Figures 3 and \@ref(fig:aspat-perf-fig-simp)).

@@

Profiling with the `Rprof` function shows which parts of a function take the longest. On a test run with 1 million cells in the weight raster and 10 landings we profiled the performance of the ILCP and MST methods. Both methods begin by constructing a full graph, which took ~ 4 s. For the ILCP method, iteratively determining the shortest paths is the only other step that takes more that a second to run, and this run time will increase with the number of landings. For the MST method, calculating cost distances among landings and from each landing to the nearest point on the existing road distance takes the most time. Determining the locations of the paths in the minimum spanning tree also takes about 2 seconds.

@@

# interactive version
# profvis::profvis(prof_input = here::here("analysis/data/derived_data/ilcp_Rprof.out"))
```

@@

>>>>

The simple cost method for setting edge weights generally performed worse than the grade penalty and Hardy methods (compare Figures 2 and \@ref(fig:aspat-perf-fig-simp), and Figures 5 and \@ref(fig:spat-perf-fig-simp)).

@@

Profiling with the `Rprof` function shows which parts of a function take the longest. On a test run with 1 million cells in the weight raster and 10 landings we profiled the performance of the ILCP and MST methods. Both methods begin by constructing a full graph, which took 3-5 s. For the ILCP method, iteratively determining the shortest paths is the only other step that takes more than a second to run, and this run time will increase with the number of landings (Figure \@ref(fig:profiling-ilcp)). For the MST method, calculating cost distances among landings and from each landing to the nearest point on the existing road distance takes the most time, which also increases with the number of landings (Figure \@ref(fig:profiling-mst)). Determining the locations of the paths in the minimum spanning tree also takes about 6 seconds. We also considered using the `gDistance` package [@van_etten_r_2017] to build the graph but it took longer than our method (Figure \@ref(fig:profiling-ilcp-gdistance)).

@@

# interactive version
# profvis::profvis(prof_input = here::here("analysis/data/derived_data/ilcp_100_Rprof.out"))
```

@@

```{r profiling-ilcp-gdistance, fig.cap="Profiling results for the ILCP method using the gdistance package internally showing functions that take > 1 s to run on a cost raster with 1 million cells. Functions higher in the call stack are called by functions lower down.", fig.height=5}
profstat(here::here("analysis/data/derived_data/ilcp_gdistance_100_Rprof.out"))
```
